---
title: "SCENIC"
author: "Joan Abinet"
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
  header-includes:
    -|
    ```{=latex}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      breaksymbolleft={}, 
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
    ```
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 45), tidy = TRUE)
```

# Load packages
```{r}
suppressMessages(library(dplyr))
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(ggplot2))
suppressMessages(library(SCENIC)) 
suppressMessages(library(ComplexHeatmap))
suppressMessages(library(AUCell))
```


```{r}
BAL_Hive.integrated <- readRDS("../Hive-2-Annotation_cells/BAL_hive.annotated_noDB.rds")
col <- c("#E2B80C", "#998025", "#AE8052", "#e6641e", "red", "#0F8140", "#4F2569", "#ABD6E4", "#1CB7D5", "#6B7BBA","#FFC0CB", "#BDBCBC", "#DF15AE")

DimPlot(BAL_Hive.integrated, cols = col)
```

# prepare cellinfo et expression matrix
```{r}
Idents(BAL_Hive.integrated) <- "Annotation"
exprMat <- BAL_Hive.integrated@assays$RNA@data
cellInfo <- data.frame(seuratCluster=Idents(BAL_Hive.integrated))
cellInfo$Annotation <- BAL_Hive.integrated$Annotation
cellInfo$nGene <- colSums(exprMat>0)
```

```{r, eval = FALSE}
dir.create ("int")
saveRDS(exprMat, file = "int/exprMat.Rds")
saveRDS(cellInfo, file = "int/cellInfo.Rds")

cellInfo <- readRDS(file = "int/cellInfo.Rds")
```


# Download binding motifs for human
```{r, eval = FALSE}
dbFiles <- c("https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-500bp-upstream-7species.mc9nr.feather",
"https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-tss-centered-10kb-7species.mc9nr.feather")

dir.create("cisTarget_databases");
setwd("cisTarget_databases") 
for(featherURL in dbFiles)
{
  download.file(featherURL, destfile=basename(featherURL)) # saved in current dir
}
```

# Initialise Scenic
```{r}
data(list="motifAnnotations_hgnc_v9", package="RcisTarget")
motifAnnotations_hgnc <- motifAnnotations_hgnc_v9

org <- "hgnc"
dbDir <- "cisTarget_databases"
dbDir <- path.expand(dbDir)
myDatasetTitle <- "SCENIC Analysis"
data(defaultDbNames)
dbs <- defaultDbNames[[org]]
scenicOptions <- initializeScenic(org=org, dbDir=dbDir, dbs=dbs, nCores=25) 
```

# Gene filtering
1. Filter by the total number of reads per gene. Keeps only the genes with at least 6 UMI counts across all samples.

2. Filter by the number of cells in which the gene is detected. 
```{r, eval = FALSE}
exprMat <- as.matrix(exprMat) # need a regulat matrix
genesKept <- geneFiltering(exprMat, scenicOptions=scenicOptions,
                           minCountsPerGene=3*.01*ncol(exprMat),
                           minSamples=ncol(exprMat)*.01)

dim(exprMat) 
exprMat_filtered <- exprMat[genesKept, ]
dim(exprMat_filtered) 
```

# Correlation
```{r, eval = FALSE}
runCorrelation(exprMat_filtered, scenicOptions)
```

```{r, eval = FALSE}
runGenie3(exprMat_filtered, scenicOptions)
```

```{r, eval = FALSE}
scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores <- 10
scenicOptions@settings$seed <- 123

runSCENIC_1_coexNetwork2modules(scenicOptions)
runSCENIC_2_createRegulons(scenicOptions)
runSCENIC_3_scoreCells(scenicOptions, exprMat_filtered)
```

# Plotting the results

## Viewing markers based on specificity score
```{r}
regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC") # require file int/3.4_regulonAUC.Rds


rss <- calcRSS(AUC=getAUC(regulonAUC), cellAnnotation=cellInfo[colnames(regulonAUC), "seuratCluster"])
rssPlot <- plotRSS(rss)
plotly::ggplotly(rssPlot$plot)
```

## Plot on Heatmap the 10 markers the most specific of each cell type
```{r, message=FALSE, warning=FALSE, results= "hide"}
rss <- rss[,c(2,3,1,4,9,8,5,11,7,10,13,6,12)]

top_tf <- c()
for (i in 1:length(colnames(rss))) {
  
  TFs <- sort(rss[,i], decreasing = T)
  top_10 <- head(TFs, 10)
  top_tf <- c(top_tf,top_10)
}

length(top_tf)

top_tf
 
# Creation of the Heatmap

regulonAUC.mat <- regulonAUC@assays@data@listData$AUC
Subset_regulonActivity <-regulonAUC.mat[names(top_tf),]

regulonActivity_byCellType_Scaled <- t(scale(t(Subset_regulonActivity), center = T, scale=T))

df <- as.data.frame(BAL_Hive.integrated$Annotation)
colnames(df) <- "CellType"

color_df <- list(CellType = c("C1 - AM" = "#E2B80C", 
                              "C2 - AM" = "#998025", 
                              "C3 - AM" = "#AE8052", 
                              "C4 - AM" = "#e6641e",
                              "C5 - AM" = "red",
                              "C6 - Mafb+ MoMac" = "#0F8140", 
                              "C7 - Neu" = "#4F2569",
                              "C8 - T cell" = "#ABD6E4", 
                              "C9 - T cell" = "#1CB7D5", 
                              "C10 - Cycling cell" = "#2D368B", 
                              "C11 - Eos" = "#FFC0CB",
                              "C12 - Epithelial cell" = "#BDBCBC",
                              "C13 - Mast cell" = "#DF15AE"))

Heatmap <- Heatmap(regulonActivity_byCellType_Scaled, name="Regulon activity", show_column_names = FALSE, 
        column_split = factor(BAL_Hive.integrated$Annotation),
        cluster_column_slices = F,
        cluster_rows = F,
        column_title_rot = 90,
        top_annotation = HeatmapAnnotation(df = df, col = color_df),
        show_column_dend = F)

#tidyHeatmap::save_pdf(Heatmap,"Scenic_Hive.pdf", width = 40, height = 50, units = "cm")
```

```{r fig.width=10, fig.height=20}
Heatmap
```


```{r}
sessionInfo()
```

