---
title: "SCENIC"
author: "Abinet Joan"
date: "`r format (Sys.time(), format='%Y-%m-%d %H:%M:%S %z')`"
output:
  pdf_document: 
    toc: true
    toc_depth: 2
  html_document: default
header-includes:
  - |
    ```{=latex}
    \usepackage{fvextra}
    \DefineVerbatimEnvironment{Highlighting}{Verbatim}{
      breaksymbolleft={}, 
      showspaces = false,
      showtabs = false,
      breaklines,
      commandchars=\\\{\}
    }
    ```
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages
```{r}
suppressMessages(library(dplyr))
suppressMessages(library(Seurat))
suppressMessages(library(patchwork))
suppressMessages(library(ggplot2))
suppressMessages(library(SCENIC)) 
suppressMessages(library(ComplexHeatmap))
suppressMessages(library(AUCell))
suppressMessages(library(formatR))
```

# Loading data
```{r }
Cell_10xAnnotate <- readRDS("../2-Paper/lavage_10x_annotated.rds")
```


# prepare cellinfo et expression matrix
```{r, eval = FALSE}
exprMat <- Cell_10xAnnotate@assays$RNA@data
cellInfo <- data.frame(seuratCluster=Idents(Cell_10xAnnotate))
cellInfo$Annotation <- Cell_10xAnnotate$Annotation
cellInfo$nGene <- colSums(exprMat>0)

dir.create ("int")
saveRDS(exprMat, file = "int/exprMat.Rds")
saveRDS(cellInfo, file = "int/cellInfo.Rds")
```

```{r}
exprMat <- readRDS("int/exprMat.Rds")
cellInfo <- readRDS("int/cellInfo.Rds")
```


# Download binding motifs for HUMAN
```{r, eval = FALSE}
dbFiles <- c("https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-500bp-upstream-7species.mc9nr.feather",
"https://resources.aertslab.org/cistarget/databases/old/homo_sapiens/hg19/refseq_r45/mc9nr/gene_based/hg19-tss-centered-10kb-7species.mc9nr.feather")

dir.create("cisTarget_databases");
setwd("cisTarget_databases") 
for(featherURL in dbFiles)
{
  download.file(featherURL, destfile=basename(featherURL)) # saved in current dir
}
```

# Initialise Scenic
```{r}
data(list="motifAnnotations_hgnc_v9", package="RcisTarget")
motifAnnotations_hgnc <- motifAnnotations_hgnc_v9

org <- "hgnc"
dbDir <- "cisTarget_databases"
dbDir <- path.expand(dbDir)
myDatasetTitle <- "SCENIC Analysis"
data(defaultDbNames)
dbs <- defaultDbNames[[org]]
scenicOptions <- initializeScenic(org=org, dbDir=dbDir, dbs=dbs, nCores=25) 
```

# Gene filtering
1. Filter by the total number of reads per gene. Keeps only the genes with at least 6 UMI counts across all samples.

2. Filter by the number of cells in which the gene is detected. 
```{r, eval = FALSE}
exprMat <- as.matrix(exprMat) # need a regulat matrix
genesKept <- geneFiltering(exprMat, scenicOptions=scenicOptions,
                           minCountsPerGene=3*.01*ncol(exprMat),
                           minSamples=ncol(exprMat)*.01)

dim(exprMat) # 14386  
exprMat_filtered <- exprMat[genesKept, ]
dim(exprMat_filtered) # 9389 
```

# Correlation
```{r, eval = FALSE}
runCorrelation(exprMat_filtered, scenicOptions)
```

```{r, eval = FALSE}
runGenie3(exprMat_filtered, scenicOptions)
```

```{r, eval = FALSE}
scenicOptions@settings$verbose <- TRUE
scenicOptions@settings$nCores <- 10
scenicOptions@settings$seed <- 123

runSCENIC_1_coexNetwork2modules(scenicOptions)
runSCENIC_2_createRegulons(scenicOptions)
runSCENIC_3_scoreCells(scenicOptions, exprMat_filtered)
```

# Plotting the results

## Viewing markers based on specificity score
```{r, message=FALSE, warning=FALSE, results= "hide", fig.show='hide'}
regulonAUC <- loadInt(scenicOptions, "aucell_regulonAUC") # require file int/3.4_regulonAUC.Rds


rss <- calcRSS(AUC=getAUC(regulonAUC), cellAnnotation=cellInfo[colnames(regulonAUC), "seuratCluster"])
#rssPlot <- plotRSS(rss["Mafb (44g)",])
rssPlot <- plotRSS(rss)
plotly::ggplotly(rssPlot$plot)
```

## Plot on Heatmap the 10 markers the most specific of each cell type
```{r, message=FALSE, warning=FALSE, results= "hide"}
#rss <- rss[,c(8,11,2,3,7,1,12,9,4,6,5,10)]
rss <- rss[,c(1,6,8,3,7,2,13,12,5,4,9,11,10)]

rss <- rss[,c(1,7,13,4,6,11,8,3,9,2,5,12,10)]

top_tf <- c()
for (i in 1:length(colnames(rss))) {
  
  TFs <- sort(rss[,i], decreasing = T)
  top_10 <- head(TFs, 10)
  top_tf <- c(top_tf,top_10)
}

length(top_tf)

#top_tf
 
# Creation of the Heatmap

regulonAUC.mat <- regulonAUC@assays@data@listData$AUC
Subset_regulonActivity <-regulonAUC.mat[names(top_tf),]

regulonActivity_byCellType_Scaled <- t(scale(t(Subset_regulonActivity), center = T, scale=T))

colors <- c("#E3751C", "#3218b5", "#52fae9", "#33A02C", "#0fc0db", "#8c6a94", "#78550b", "lightblue", "#7486fc", "grey", "#600078", "gray", "black")

names(colors) <- levels(Cell_10xAnnotate$Annotation) 

df <- as.data.frame(Cell_10xAnnotate$Annotation)
colnames(df) <- "CellType"

color_df <- list(CellType = c("C1 - AM" = "#E3751C", 
                              "C2 - AM" = "#78550b", 
                              "C3 - AM" = "black", 
                              "C4 - MoMac" = "#33A02C", 
                              "C5 - Neu" = "#8c6a94", 
                              "C6 - Neu" = "#600078", 
                              "C7 - LT" = "lightblue", 
                              "C8 - LT" = "#52fae9", 
                              "C9 - LT" = "#7486fc", 
                              "C10 - LT" = "#3218b5", 
                              "C11 - LT" = "#0fc0db",
                              "C12 - Mast" = "grey",
                              "C13 - Epith" = "grey"))

Heatmap <- Heatmap(regulonActivity_byCellType_Scaled, name="Regulon activity", show_column_names = FALSE, 
        column_split = factor(Cell_10xAnnotate$Annotation),
        cluster_column_slices = F,
        cluster_rows = F,
        top_annotation = HeatmapAnnotation(df = df, col = color_df),
        show_column_dend = F,
        column_title_rot = 90)

#tidyHeatmap::save_pdf(Heatmap,"Scenic_10x.pdf", width = 40, height = 50, units = "cm")
```

```{r fig.width=10, fig.height=20}
Heatmap
```


```{r}
sessionInfo()
```

